An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-10-02 paper.tex
Update 5.3.5

An important question~\cite{li2021does,guo2020inclusive} for any
organisation is whether higher level participants encourage those at lower
levels in the hierarchy to properly engage. The IETF is cooperative and
voluntary by design, therefore we hypothesise a high level of encouragement
from higher levels to lower. We measure whether activity in one time period
is associated with activity in a subsequent time period both for
individuals and for neighbourhoods. If those who communicate with higher
levels receive a boost in their communication as a result, this is an
indication of such encouragement.

In particular, we perform Mobility Taxonomy analysis to determine the
effect individual IETF participants have on their future number of
connections, and the average number of connections of their neighbours.  We
find that WGCs have an increased tendency, versus RPs, to remain active
(\textit{Mobility}), an increased tendency for their neighbours to be
active subsequent to high WGC activity (\textit{Philanthropy}) and for them
to gain activity after their neighbours are active (\textit{Community}).
We interpret this as WGCs filling a ``facilitator" role on the mailing
lists.

@@

Finally, to determine how individuals and their neighbours affect each
other's activity levels in subsequent time periods, for \textbf{RQ4}, we
take correlations of degree between different time windows using the
Mobility Taxonomy~\cite{barnes2023measuring} aspects called
\textit{Mobility}, \textit{Neighbour Mobility}, \textit{Philanthropy} and
\textit{Community} explained in the previous section. In
Figure~\ref{fig:mob_tax}, the time window chosen is one year which is split
into two snapshots graphs of six months, the time window then moves forward
one month and the process is repeated. The mid point of the year time
window is plotted. WGCs and RPs are plotted but the small number of ADs
mean that the correlations are too noisy to show meaningful effects.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\figureWidthTwoColumn]{figures-prev/icwsm-2024/hierarchy_status_comparison.png}
  \caption{
    Connection between roles and activity for RP and WGC viewed in terms of
    a) Mobility b) Neighbour Mobility c) Philanthropy and d) Community over
    time. See section~\ref{subsec:mob_tax} for definitions of these terms.
  }
  \label{fig:mob_tax}
\end{figure*}

The \textit{Mobility} of both RPs and WGCs is plotted in
Figure~\ref{fig:mob_tax}a. The high amount of correlation between the
degree of an individual in the first half of the time window with the
second half shows that all users have a tendency to become more active in
subsequent time periods if they are highly active in the previous period,
but this is more pronounced in WGCs. This is found to follow a period of
about one year for both RPs and WGC, suggesting that a person is likely to
be as active as there were a year ago. The \textit{Neighbour Mobility}
plot, Figure~\ref{fig:mob_tax}b, shows a similar tendency for
neighbourhoods, but with less of a split.

The \textit{Philanthropy} plot, Figure~\ref{fig:mob_tax}c, shows a higher
correlation for WGCs than RPs between an individual's degree in the first
snapshot with and their ND in the second. This means that individuals who
directly interact with WGCs become more active on mailing lists
subsequently. The \textit{Community} plot, Figure~\ref{fig:mob_tax}d, shows
the correlation between the ND in the first snapshot with the degree in the
second.  The higher correlation for WGCs suggests they benefit from
interactions with highly active neighbours by themselves becoming more
active in subsequent time periods. These trends for \textit{Philanthropy}
and \textit{Community} show a reversal in early 2016 that coincides with a
period where the mean number of WGCs per WG and the number of WGC roles was
rapidly rising, see figure~\ref{fig:wgc_over_time}. A merging of the
Applications area and Real-time Applications \& Infrastructure area in May
2015 may be a cause of this reversal of correlation, as the assignment of
WGs to areas was changed which may cause more non-reciprocal WGC
communication than normal.

Increased \textit{Mobility} for WGCs suggests that they gain advantage from
their role with their activity leading to similar communication activity in
the future. Moreover, they experience an increased effect of
\textit{Philanthropy} and \textit{Community}. The former suggests that if a
WGC is engaged in a higher amount of discussion, then in the future the
individuals who they were communicating with will also be highly active.
The latter suggests that when WGCs are surrounded by people who are active
in discussion, they are likely to be more active themselves in the future
too. All of this has less of an effect for RPs who show a lower correlation
in terms of all three aspects. Our interpretation of this is that active
WGCs are ``facilitating" discussion in the WG mailing lists; both having
their activity boosted by their neighbours and, in turn, boosting the
activities of their neighbours. This aligns well with the hypothesis that
the cooperative and voluntary design of the IETF lends itself well to
higher levels encouraging those lower to engage in discussion.

Overall the answer to RQ4 is that the WGCs have a positive effect on their
immediate contacts and those individuals that directly discuss with WGCs
are more likely to engage more in discussion in subsequent time periods.
Conversely though the WGCs who discuss topics with individuals who are
active in discussion are, themselves, more likely to engage more fully in
discussion. This points to something like a virtuous cycle of WGCs
encouraging and being encouraged by their direct contacts.

>>>>

An important question for any organisation is whether higher level
participants encourage those at lower levels in the hierarchy to properly
engage \cite{li2021does,guo2020inclusive}. The IETF is cooperative and
voluntary by design, therefore we hypothesise a high level of encouragement
from higher levels to lower. We measure whether activity in one time period
is associated with activity in a subsequent time period both for
individuals and for neighbourhoods. If those who communicate with higher
levels receive a boost in their communication as a result, this is an
indication of such encouragement.

% In particular, we perform Mobility Taxonomy analysis to determine the
% effect individual IETF participants have on their future number of
% connections, and the average number of connections of their neighbours.  We
% find that working group chairs have an increased tendency, versus regular
% participants, to remain active (\textit{Mobility}), an increased tendency
% for their neighbours to be active subsequent to high working group chair
% activity (\textit{Philanthropy}) and for them to gain activity after their
% neighbours are active (\textit{Community}).  We interpret this as working
% group chairs filling a ``facilitator'' role on the mailing lists.

%..................................................................................................
% From our ICWSM 2024 paper, Section 4.3:

\pb{Mobility Taxonomy:}
We aim to measure the association of the activity of nodes, and their
neighbours, with their activity in a subsequent time period. To determine
this association we use the temporal graph analysis technique called the
Mobility Taxonomy~\cite{barnes2023measuring}, which involves correlating
the degree and average neighbourhood degree (ND) between two adjoining time
windows, both half of the chosen time window length for other analysis. The
temporal graph analysis tool Raphtory~\cite{steer2020raphtory} is used to
calculate the raw degree and ND numbers before performing these
correlations.

There are six combinations to correlate node degree and ND in time window
one and two, the combinations are outlined in detail
in~\cite{barnes2023measuring}. The four measures used in this paper are
\textit{Mobility}, \textit{Neighbour Mobility}, \textit{Philanthropy} and
\textit{Community}.  Each represents the correlation of the degree of a
node (or set of nodes) at an earlier or later period of time.
\textit{Mobility} can be thought of as the tendency for a node that is
active in the first time window to be equally as active in the second.
\textit{Neighbour Mobility} is similar but for a node's neighbourhood.
\textit{Philanthropy} can be thought of as the tendency of a node's
neighbourhood to be active in window two if that node is active in window
one. Finally, \textit{Community} is the reverse of \textit{Philanthropy} as
the activity of a node in window two is compared to that node's neighbours
activity in window one.

Only the nodes which are active in the first window are considered for the
second window. Nodes inactive in the second time window are said to have a
degree of 0. The ND in the second time window is taken over the same set of
neighbours that existed in the first; that is we look at a node's
neighbours in the first window and measure their degree in the second for
consistency of comparison.

@@

\pb{Participant Mobility:}
To determine how individuals and their neighbours affect each other's
activity levels in subsequent time periods, we take correlations of degree
between different time windows using this Mobility Taxonomy. 
Figure~\ref{fig:mob_tax}, the time window chosen is one year which is split
into two snapshots graphs of six months, the time window then moves forward
one month and the process is repeated. The mid point of the year time
window is plotted. Working group chairs and regular participants are
plotted but the small number of area directors mean that the correlations
are too noisy to show meaningful effects.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\figureWidthTwoColumn]{figures-prev/icwsm-2024/hierarchy_status_comparison.png}
  \caption{
    Connection between roles and activity for regular participants and
    working group chairs viewed in terms of a) Mobility b) Neighbour
    Mobility c) Philanthropy and d) Community over time.
  }
  \label{fig:mob_tax}
\end{figure*}

The \textit{Mobility} of regular participants and working group chairs is
plotted in Figure~\ref{fig:mob_tax}(a). The high amount of correlation
between the degree of an individual in the first half of the time window
with the second half shows that all users have a tendency to become more
active in subsequent time periods if they are highly active in the previous
period, but this is more pronounced in working group chairs. This is found
to follow a period of about one year for both regular participants and
working group chairs, suggesting that a person is likely to be as active as
there were a year ago. The \textit{Neighbour Mobility} plot,
Figure~\ref{fig:mob_tax}(b), shows a similar tendency for neighbourhoods, but
with less of a split.

The \textit{Philanthropy} plot, Figure~\ref{fig:mob_tax}(c), shows a higher
correlation for working group chairs than regular participants between an
individual's degree in the first snapshot with and their ND in the second.
This means that individuals who directly interact with working group chairs
become more active on mailing lists subsequently. The \textit{Community}
plot, Figure~\ref{fig:mob_tax}(d), shows the correlation between the ND in
the first snapshot with the degree in the second.  The higher correlation
for working group chairs suggests they benefit from interactions with
highly active neighbours by themselves becoming more active in subsequent
time periods. These trends for \textit{Philanthropy} and \textit{Community}
show a reversal in early 2016 that coincides with a period where the mean
number of chairs per working group, and the number of working group chairs,
rapidly rising, see Figure~\ref{fig:wgc_over_time}. A merging of the
Applications area and Real-time Applications \& Infrastructure area in May
2015 may be a cause of this reversal of correlation, as the assignment of
working group chairs to areas was changed which may cause more non-reciprocal 
communication than normal.

\pb{Summary:}
Increased \textit{Mobility} for working group chairs suggests that they
gain advantage from their role with their activity leading to similar
communication activity in the future. Moreover, they experience an
increased effect of \textit{Philanthropy} and \textit{Community}. The
former suggests that if a working group chair is engaged in a higher amount
of discussion, then in the future the individuals who they were
communicating with will also be highly active.  The latter suggests that
when working group chairs are surrounded by people who are active in
discussion, they are likely to be more active themselves in the future too.
All of this has less of an effect for regular participants who show a lower
correlation in terms of all three aspects. Our interpretation of this is
that active working group chairs are ``facilitating'' discussion in the WG
mailing lists; both having their activity boosted by their neighbours and,
in turn, boosting the activities of their neighbours. This aligns well with
the hypothesis that the cooperative and voluntary design of the IETF lends
itself well to higher levels encouraging those lower to engage in
discussion.

Working group chairs have a positive effect on their immediate contacts.
Those individuals that directly discuss with working group chairs are more
likely to engage more in discussion in subsequent time periods.  Conversely
though the working group chairs who discuss topics with individuals who are
active in discussion are, themselves, more likely to engage more fully in
discussion. This points to something like a virtuous cycle of chairs
encouraging and being encouraged by their direct contacts.
